### 2021-10-15 plos_latex_template.tex
Rocío: editing rule language

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, races, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for gaps in its composition.

%ast: I added race -even if race is a construct- because focus is given to gender and the spaces tend to be predominantly white and US/Europe-centric. In other parts of the text I was using racialized but here we can keep it simple, i think. another thing: i avoided "gender balancing" because idk if that comes across as too binary, even if what is done is adding cis women to the team
% Rocío: what about ethnicity instead of race? I agree with the gender balancing thing.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}. Excluding them may potentially lead to missing innovative contributions due to a language barrier.

% Liz: People with disabilities don't want to be considered "needy". We reject the term "special needs" in part because it frames the person with the disability as being the burden causing the accommodation, where maybe it's the responsibility of society to be open to everyone. So I wonder if it's the same for people who speak other languages. Maybe they are not needy or inferior.  Or maybe it's wrong for me to apply disability rights philosophy to language.
The linguistic diversity of conference participants is often overlooked, resulting in missed opportunities for them, as well as for the conference to benefit from their contributions. Non-native English speakers may feel intimidated to ask questions or raise their views. Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2). If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.

Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

% Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)

>>>>

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, races and ethniticities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for gaps in its composition.

% Rocío: what about ethnicity instead of race? %i am not sure

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international conferences, the linguistic diversity of conference participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
and conferences may miss potentially innovative contributions.

% Rocío: main idea: make conferences more linguistically inclusive
Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (Rule 2). 
Advertising the conference in several languages and considering having non-English workshops and presentations to help overcome this barrier. 
Allow the audience to ask questions in their native language and invite others to help translate it; conversely, if the presenter is not comfortable with questions in English, encourage people in the audience to help by rephrasing their questions or try to translate them; after all, the community should be supportive.

% Not citing ninerBetterWhomLeveling2021 anymore for now.

@@

% % Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
% As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
